\documentclass{beamer}
%Information to be included in the title page:
\title[ANTIQUE Seminar]{ANTIQUE Seminar - Relational Symbolic Execution}
\author{Ignacio Tiraboschi}
\institute[INRIA / ENS]{INRIA Paris - ENS Paris}
\date{2022}
\usetheme{CambridgeUS}
\setbeamertemplate{page number in head/foot}[appendixframenumber]

% if then else in macros
\usepackage{xparse}
% inference rules
\usepackage{mathpartir}
% math fonts
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{mathtools}
\usepackage{mathrsfs}
\usepackage{latexsym}
%
\usepackage{tikz}
\usepackage{minted}
\usepackage{amsmath}
\usepackage{caption}
\usepackage{subcaption}
\usepackage[round,semicolon,authoryear,sort]{natbib}
\bibliographystyle{unsrtnat}
%
\RecustomVerbatimEnvironment{Verbatim}{BVerbatim}{}
\renewcommand{\figurename}{Listing}
%
\newcommand{\TODO}[1]{\noindent \textcolor{red}{\textsf{TODO: #1}}}
%
\input{macros.tex}
%
\begin{document}

\frame{\titlepage}

\begin{frame}
  \frametitle{Outline}
  \begin{enumerate}
    \item Symbolic execution.
    \begin{itemize}
      \item What is Symbolic execution?
      \item Example.
      \item Pros and cons.
    \end{itemize}
    \item Hyperproperties.
    \begin{itemize}
      \item Noninterference.
    \end{itemize}
    \item Symbolic execution for noninterference: \emph{Relational symbolic execution}.
    \begin{itemize}
      \item Example.
      \item How to achieve soundness?
    \end{itemize}
    \item 
  \end{enumerate}
\end{frame}

\begin{frame}
  \frametitle{Symbolic execution}
  \emph{Symbolic execution} executes the program over \textbf{symbolic variables and values} which
  \textbf{represent possible concrete values}.
  
  \pause
  \vspace{1em}
  Symbolic execution is very expressive and precise, and this is particularly good to find
  security vulnerabilities.
  
  \pause
  \vspace{1em}
  It is only \emph{sound up to a bound}. We can only unroll loops a limited amount of times.

  \pause
  \vspace{1em}
  It relies on an off-the-shelf SMT solver to handle satisfiabilty queries, which can eventually
  be a problem when queries are too complicated. Also, if not careful, there is
  \textbf{path explosion} \citep{DBLP:journals/cacm/CadarS13}.
\end{frame}

\begin{frame}
  \frametitle{Symbolic execution}
  More precisely, symbolic state is a pair $(\rho,\pi)$ where $\rho$ is the symbolic mapping and
  $\pi$ is the symbolic path.

  \vspace{1em}
  A regular safety property could be ``the program never divides by zero'' .
  To check this we need to see that:
  \begin{center}
    In all possible traces, no trace ever divides by zero.
  \end{center}

  \vspace{1em}
  Let us see an example to understand how we can use symbolic execution for verification.
\end{frame}

\begin{frame}[fragile]
  \frametitle{Example}
  \begin{figure}[htp]
    \centering
    \begin{minted}{c}
int x, y;
if (x != 0) {
  y = y / x;
}
    \end{minted}
  \end{figure}
  \begin{figure}
    \begin{tikzpicture}
      \node[scale=0.8] { $\rho = \{ \textsf{x} \rightarrow x_0\ ;\ \textsf{y} \rightarrow y_0 \} \;
              \pi = \text{True}$ } [sibling distance = 5.5cm]
        child [red] {
          node [scale=0.8] {
            $\rho = \{ \textsf{x} \rightarrow x_0\ ;\ \textsf{y} \rightarrow y_0 \} \;
            \pi = \{ x_0 = 0 \}$
            } edge from parent node [scale=0.8, left, brown] {\texttt{if-false}}}
        child [blue] {
          node [scale=0.8] {
            $\rho = \{ \textsf{x} \rightarrow x_0\ ;\ \textsf{y} \rightarrow y_0 \} \;
            \pi = \{ x_0 \not = 0 \}$
            }
            child {node [scale=0.8] {
              $\rho = \{ \textsf{x} \rightarrow x_0\ ;\ \textsf{y} \rightarrow y_0 / x_0 \} \;
              \pi = \{ x_0 \not = 0 \}$
              } edge from parent node [scale=0.8, right, brown] {\texttt{assign}}}
            edge from parent node [scale=0.8, right, brown] {\texttt{if-true}}};
    \end{tikzpicture}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Symbolic execution}
  The previous example works great... \pause but what if we had loops?
  
  \vspace{1em}
  The usual method to deal with loops is to unfold up to certain bound, losing capability to actually
  verify a property.
  But this still allows finding vulnerabilities in the program.
\end{frame}

\begin{frame}
  \frametitle{Noninterference}
  The previous property depends only on individual traces.

  \vspace{1em}
  A \textbf{2-safety} \citep{Clarkson2008} property is a safety property over sets of traces.
  One 2-safety property of interest is \emph{noninterference}.

  \pause
  \vspace{1em}
  Noninterference is a confidentiality policy that stipulates commands executed on behalf of
  users holding \textbf{high clearances have no effect} on system behavior observed by users holding
  \textbf{low clearances}.

  \pause
  \begin{center}
    For any two traces initially agreeing on public values, if the traces end, they agree on its
    public values.
  \end{center}
\end{frame}

\begin{frame}
  \frametitle{Relational symbolic execution}
  In \emph{relational symbolic execution} \citep*{DBLP:journals/corr/abs-1711-08349} the authors
  propose a framework for verification of properties such as noninterference.
  
  \pause
  \vspace{1em}
  In relational symbolic execution we have that each variable is mapped to either one or two
  symbolic expressions.
  \[
    \textsf{x} \rightarrow \mpair{x_0}{x_1} \qquad \textsf{y} \rightarrow \msingle{y_0}
  \]

  \pause
  This allows for optimization:
  \begin{itemize}
    \item Low clearance variables cannot branch leading to fewer symbolic paths to follow.
    \item This also means that we execute fewer queries to the SMT solver.
  \end{itemize}
  \centering
  \textbf{Now we also have many possible combinations of paths!}
\end{frame}

\begin{frame}[fragile]
  \frametitle{Example}
  \begin{figure}[htp]
  \centering
  \begin{minted}{c}
if (x > 0) {
  y = 10;
}
  \end{minted}
  \end{figure}
  \begin{figure}
    \begin{tikzpicture}
      \node[scale=0.8] { $\rho = \{ \textsf{x} \rightarrow \mpair{x_0}{x_1}\ ;\ \textsf{y} \rightarrow \msingle{y_0} \} \;\;
              \pi = \text{True}$ } [sibling distance = 4.5cm]
        child [red] {
          node [scale=0.8] { \vdots } edge from parent node [scale=0.8, left, brown] {\texttt{if-false-true}}}
        child [blue] {
          node [scale=0.8] {
            $\rho = \{ \textsf{x} \rightarrow \mpair{x_0}{x_1}\ ;\ \textsf{y} \rightarrow \msingle{y_0} \} \;\;
            \pi = \{ x_0 > 0\ \wedge\ x_1 \leq 0 \}$
            }
            child {node [scale=0.8] {
              $\rho = \{ \textsf{x} \rightarrow \mpair{x_0}{x_1}\ ;\ \textsf{y} \rightarrow \mpair{10}{y_0} \} \;\;
              \pi = \{ x_0 > 0\ \wedge\ x_1 \leq 0 \}$
              } edge from parent node [scale=0.8, right, brown] {\texttt{assign}}}
            edge from parent node [scale=0.8, right, brown] {\texttt{if-true-false}}};
    \end{tikzpicture}
  \end{figure}
\end{frame}

\begin{frame}
  \frametitle{Vulnerabilities in Relational symbolic execution}
  In the previous example, we are able to see that at least one relational trace is violating
  noninterference.

  \vspace{1em}
  Thanks to the SMT solver, we are able to get a model of showing the vulnerability. We can also
  reconstruct the path by looking at the symbolic path $\spath$.
\end{frame}

\begin{frame}
  \frametitle{Soundness in Relational symbolic execution}
  As we noted earlier, symbolic execution usually gives up on soundness and is used to find
  vulnerabilities.

  \vspace{1em}
  In the framework of Relational symbolic execution what is done is to \textbf{manually set invariants}
  to over approximate loops, and skip the symbolic execution of loops.
\end{frame}



\begin{frame}[allowframebreaks]
\frametitle{Bibliography}
\bibliography{main}
\end{frame}

\end{document}
